---
title: "Heterogeneous Treatment Effects in a Crowdsourced Megastudy of Single-Session Interventions for Depression"
subtitle: "Reproducible analysis"
date: today
date-format: "MMMM D, YYYY"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    df-print: paged
    code-fold: false
    embed-resources: true
execute:
  warning: false
  message: false
  cache: false
editor: visual
---

# Overview

Reproduces the secondary HTE analysis of a crowdsourced megastudy of single-session interventions (SSIs) for depression (N = 7,505; 12 SSIs vs. control). Run `Rscript R/01_generate_ipw_weights.R` first to produce `data/sl_weights.csv`, then render this document. Tables and figures are written to `output/`. See `README.md` for details.

# Setup

```{r setup}
library(tidyverse)
library(grf)
library(mclust)
library(glmnet)
library(tidymodels)
library(emmeans)
library(broom)
library(boot)
library(corrplot)
library(reshape2)
library(ggrepel)
library(car)

# Serpentine sort + fold helper (self-contained; no external packages)
source(file.path("R", "sl_helpers.R"))

ssi_labels <- c(
  "5 Habits to Beat Depression", "AI-assisted Personalized Narrative",
  "Behavioral Activation", "Dialectical Behavioral Therapy Skills",
  "Inner Child Healing Walk", "Interactive Cognitive Reappraisal",
  "Mindful Acceptance", "Mindful Attention Skills", "Moral Elevation",
  "Personalized Intervention Recommender", "Reframing Negative Thoughts",
  "Savoring Strategies"
)

# ---- Paths (relative to this document) ---------------------------------
DATA_DIR    <- "data"
TABLES_DIR  <- file.path("output", "tables")
FIGURES_DIR <- file.path("output", "figures")

dir.create(TABLES_DIR,  showWarnings = FALSE, recursive = TRUE)
dir.create(FIGURES_DIR, showWarnings = FALSE, recursive = TRUE)
```

# Data load and cleaning

```{r data-load}
mydata <- read.csv(file.path(DATA_DIR, "ml_clean.csv"), header = TRUE)
mydata <- mydata[, -1]  # drop leading row-number column
mydata <- mydata %>% select(-help_prefer_not_say_baseline)  # zero-variance

# Reverse-score PHQ-9 follow-up: higher = better, matches HTE convention
mydata$phq9_score_followup <- 27 - mydata$phq9_score_followup

# Build PHQ-9 baseline raw + reverse-scored versions
items_phq9_baseline    <- paste0("phq9_", 1:9, "_baseline")
mydata$phq9_pre_raw    <- rowSums(mydata[, items_phq9_baseline], na.rm = TRUE)
mydata$phq9_pre_rev    <- 27 - mydata$phq9_pre_raw

# Follow-up completion indicator
mydata$fup_complete <- as.integer(!is.na(mydata$phq9_score_followup))

# Intervention-vs-control indicator
mydata$int_binary <- ifelse(mydata$group_numeric == 13, 0, 1)

# Note: `group_numeric` is left as integer here so it joins cleanly with
# `sl_weights.csv` in the next chunk. It is converted to a factor
# (with control = 13 as reference) later, after the IPW merge.

cat("N randomized:", nrow(mydata), "\n")
cat("N completers (4-week PHQ-9):", sum(mydata$fup_complete), "\n")
```

## Drop low-variance covariates

```{r drop-low-var}
# Baseline covariate columns (positions 4:102 in the cleaned file)
W_full <- mydata[, 4:102]

constant_vars <- sapply(W_full, function(x) length(unique(x)) <= 1)
low_variance  <- sapply(W_full, function(x) {
  if (is.numeric(x) && all(x %in% c(0, 1))) {
    min(sum(x == 0, na.rm = TRUE), sum(x == 1, na.rm = TRUE)) < 5
  } else {
    FALSE
  }
})

keep_cols <- !(constant_vars | low_variance)
cat("Dropped",  sum(!keep_cols), "low/zero-variance baseline covariates\n")
cat("Retained", sum(keep_cols), "baseline covariates for CATE estimation\n")
```

# Attach loss-to-follow-up weights

Join the weights from step 1 onto `mydata` by `pid` × `group_numeric`.

```{r attach-weights}
weights_path <- file.path(DATA_DIR, "sl_weights.csv")
if (!file.exists(weights_path)) {
  stop("data/sl_weights.csv not found. Run `Rscript R/01_generate_ipw_weights.R` first.")
}

weights_long <- read.csv(weights_path)
weights_to_merge <- weights_long %>%
  select(pid, intervention, w_norm) %>%
  distinct()

mydata <- mydata %>%
  left_join(weights_to_merge,
            by = c("pid", "group_numeric" = "intervention"))

stopifnot(sum(is.na(mydata$w_norm)) == 0)

# Now safe to convert `group_numeric` to a factor with control as reference.
mydata$group_numeric <- factor(mydata$group_numeric)
mydata$group_numeric <- relevel(mydata$group_numeric, ref = "13")
```

# CATE estimation --- Generalized Causal Forests with 10-fold CV

For each SSI, a CATE relative to control (arm 13) is estimated from `grf::causal_forest` fit on 10 training folds (`make_folds_serpentine()`). Focal-arm and control participants get out-of-fold predictions; everyone else gets the average of the 10 forests.

```{r cate-estimate}
df_cate <- {

  controls   <- mydata %>% filter(group_numeric == 13)
  controls_f <- make_folds_serpentine(controls, nfolds = 10, seed = 123)
  control_folds_map <- controls_f %>% select(pid, fold)

  arms <- 1:12
  n_participants <- nrow(mydata)
  pid_vec <- mydata$pid

  mat_cate <- matrix(NA_real_,
                     nrow = n_participants,
                     ncol = length(arms),
                     dimnames = list(NULL, sprintf("CATE_%d", arms)))

  X_all <- as.matrix(mydata[, 4:102])

  for (arm in arms) {
    message(sprintf("Arm %d vs 13 (10-fold CV)", arm))

    arm_dat <- mydata %>% filter(group_numeric == arm)
    arm_f   <- make_folds_serpentine(arm_dat, nfolds = 10, seed = 1000 + arm)
    arm_folds_map <- arm_f %>% select(pid, fold)

    analytic <- mydata %>%
      filter(group_numeric %in% c(13, arm)) %>%
      left_join(control_folds_map, by = "pid") %>%
      left_join(arm_folds_map,     by = "pid",
                suffix = c("_ctrl", "_arm")) %>%
      mutate(fold = ifelse(group_numeric == 13, fold_ctrl, fold_arm)) %>%
      select(-fold_ctrl, -fold_arm)

    analytic_obs <- analytic %>% filter(!is.na(phq9_score_followup))

    oof_tau <- rep(NA_real_, nrow(analytic_obs))
    forests <- vector("list", 10)

    for (k in 1:10) {
      train_k <- analytic_obs %>% filter(fold != k)
      test_k  <- analytic_obs %>% filter(fold == k)
      if (nrow(train_k) < 20 || nrow(test_k) < 1) next

      X_train  <- as.matrix(train_k[, 4:102])
      Y_train  <- as.numeric(train_k$phq9_score_followup)
      W_train  <- as.integer(train_k$group_numeric == arm)
      wt_train <- as.numeric(train_k$w_norm)

      set.seed(2000 + arm * 100 + k)
      cf_k <- grf::causal_forest(
        X_train, Y_train, W_train,
        sample.weights = wt_train,
        num.trees      = 4000
      )
      forests[[k]] <- cf_k

      X_test <- as.matrix(test_k[, 4:102])
      idx    <- which(analytic_obs$fold == k)
      oof_tau[idx] <- as.numeric(predict(cf_k, X_test)$predictions)
    }

    tau_arm_all <- rep(NA_real_, n_participants)
    tau_arm_all[match(analytic_obs$pid, mydata$pid)] <- oof_tau

    # Predict onto everyone NOT in controls or this arm using the saved forests.
    other_idx <- which(!(mydata$group_numeric %in% c(13, arm)))
    pred_mat  <- matrix(NA_real_, nrow = length(other_idx), ncol = 10)
    for (k in 1:10) {
      if (is.null(forests[[k]])) next
      pred_mat[, k] <- as.numeric(
        predict(forests[[k]], X_all[other_idx, , drop = FALSE])$predictions)
    }
    tau_arm_all[other_idx] <- rowMeans(pred_mat, na.rm = TRUE)

    # Backfill any remaining NAs (skipped folds, missing outcomes in arm/control)
    miss_idx <- which(is.na(tau_arm_all))
    if (length(miss_idx) > 0) {
      pred_mat_all <- matrix(NA_real_, nrow = length(miss_idx), ncol = 10)
      for (k in 1:10) {
        if (is.null(forests[[k]])) next
        pred_mat_all[, k] <- as.numeric(
          predict(forests[[k]], X_all[miss_idx, , drop = FALSE])$predictions)
      }
      tau_arm_all[miss_idx] <- rowMeans(pred_mat_all, na.rm = TRUE)
    }

    mat_cate[, arm] <- tau_arm_all
  }

  data.frame(pid = pid_vec, mat_cate)
}

stopifnot(sum(is.na(df_cate)) == 0)
head(df_cate)
```

## GCF calibration → Supplementary Table 2

Per arm, fit a causal forest on the focal-arm + control sample and run `grf::test_calibration` (mean- and differential-prediction coefficients).

```{r gcf-calibration}
calibration_results <- {
  out <- data.frame()
  for (arm in 1:12) {
    analytic_obs <- mydata %>%
      filter(group_numeric %in% c(arm, 13),
             !is.na(phq9_score_followup))

    X_full  <- as.matrix(analytic_obs[, 4:102])
    Y_full  <- as.numeric(analytic_obs$phq9_score_followup)
    W_full  <- as.integer(analytic_obs$group_numeric == arm)
    wt_full <- as.numeric(analytic_obs$w_norm)

    set.seed(3000 + arm)
    cf_full <- grf::causal_forest(
      X_full, Y_full, W_full,
      sample.weights = wt_full,
      num.trees      = 4000
    )
    cal <- test_calibration(cf_full)

    out <- rbind(out, data.frame(
      arm      = arm,
      mean_est = cal[1, 1], mean_se = cal[1, 2], mean_p = cal[1, 4],
      diff_est = cal[2, 1], diff_se = cal[2, 2], diff_p = cal[2, 4]
    ))
  }
  out
}

calibration_results$SSI <- ssi_labels[calibration_results$arm]
cal_out <- calibration_results[, c("arm", "SSI", "mean_est", "mean_se", "mean_p",
                                    "diff_est", "diff_se", "diff_p")]
write.csv(cal_out,
          file.path(TABLES_DIR, "supp_table2_calibration.csv"),
          row.names = FALSE)
print(cal_out, digits = 3)
```

## CATE 12 × 12 correlation → Supplementary Figure 1

```{r cate-correlations}
cate_cor <- cor(df_cate %>% select(starts_with("CATE_")), use = "complete.obs")
colnames(cate_cor) <- rownames(cate_cor) <- paste("SSI", 1:12)

png(file.path(FIGURES_DIR, "supp_fig1_cate_correlations.png"),
    width = 7, height = 7, units = "in", res = 300)
corrplot(cate_cor, method = "color", type = "upper",
         tl.col = "black", tl.srt = 45, addCoef.col = "black",
         number.cex = 0.7,
         col = colorRampPalette(c("blue", "white", "red"))(200))
dev.off()

corrplot(cate_cor, method = "color", type = "upper",
         tl.col = "black", tl.srt = 45, addCoef.col = "black",
         number.cex = 0.7,
         col = colorRampPalette(c("blue", "white", "red"))(200))
```

# Latent profile analysis (LPA) on CATE matrix

Standardize the 12 CATE columns and fit `mclust::Mclust(G = 1:10)` across 10 seeds; keep the best-BIC run as `mod_train`.

```{r lpa}
X_train <- scale(as.matrix(df_cate %>% select(-pid)))

mod_train <- {
  n_runs <- 10
  models <- vector("list", n_runs)
  for (i in 1:n_runs) {
    set.seed(123 + i)
    models[[i]] <- Mclust(X_train, G = 1:10)
  }
  bics <- sapply(models, function(m) m$bic)
  best <- which.max(bics)
  models[[best]]
}

cat("Best LPA: G =", mod_train$G,
    " modelName =", mod_train$modelName,
    " BIC =", round(mod_train$bic, 1), "\n")

# Compare solutions across G and covariance parameterizations
cat("Top models by BIC (model, G):\n")
print(summary(mod_train$BIC))
cat("\nFull BIC matrix (rows = G, cols = covariance model):\n")
print(round(mod_train$BIC, 1))

cl_train <- mod_train$classification
df_cate$cluster_assignment <- cl_train

mydata <- mydata %>%
  left_join(df_cate %>% select(pid, cluster_assignment), by = "pid")

cat("\nCluster sizes:\n")
print(table(mydata$cluster_assignment))
```

## Figure 1 --- CATE profiles by cluster

```{r figure1}
cluster_means_df <- as.data.frame(mod_train$parameters$mean)
colnames(cluster_means_df) <- paste("Cluster", seq_len(mod_train$G))
cluster_means_df_t <- t(cluster_means_df)
cluster_means_long <- reshape2::melt(cluster_means_df_t)
colnames(cluster_means_long) <- c("Cluster", "Variable", "Mean")
cluster_means_long$Cluster <- factor(cluster_means_long$Cluster)

last_points <- cluster_means_long %>%
  group_by(Variable) %>%
  filter(Cluster == last(levels(Cluster)))

fig1 <- ggplot(cluster_means_long,
               aes(x = Cluster, y = Mean, color = Variable, group = Variable)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_text_repel(
    data = last_points,
    aes(label = gsub("CATE_", "SSI ", Variable)),
    nudge_x = 0.3, direction = "y", hjust = 0,
    segment.color = "grey70", size = 3
  ) +
  labs(x = "Cluster",
       y = "Mean CATE (standardized across participants)") +
  scale_color_viridis_d(option = "turbo") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",
        plot.margin = margin(10, 60, 10, 10))

ggsave(file.path(FIGURES_DIR, "cate_profiles.png"),
       fig1, width = 9, height = 6, dpi = 300)

fig1
```

## Supplementary Table 5 --- Mean CATEs by SSI and cluster

```{r mean-cate-table}
mean_cate_tbl <- as.data.frame(mod_train$parameters$mean) %>%
  setNames(paste("Cluster", seq_len(mod_train$G))) %>%
  tibble::rownames_to_column("ssi") %>%
  mutate(
    ssi_num = as.integer(str_extract(ssi, "\\d+")),
    No.     = ssi_num,
    SSI     = ssi_labels[ssi_num],
    Target  = case_when(
      ssi_num %in% c(1, 5, 7, 8, 9) ~ "Cluster 1",
      ssi_num %in% c(10, 12)         ~ "Cluster 3",
      TRUE                           ~ ""
    )
  ) %>%
  arrange(ssi_num) %>%
  mutate(across(starts_with("Cluster "), \(x) round(x, 2))) %>%
  select(No., SSI, Target,
         `Cluster 1`, `Cluster 2`, `Cluster 3`, `Cluster 4`)

print(mean_cate_tbl)
write.csv(mean_cate_tbl,
          file.path(TABLES_DIR, "supp_table5_mean_cates.csv"),
          row.names = FALSE)
```

## Supplementary Table 3 --- Classification quality

Per-cluster mean modal posterior probability and relative entropy ($1 - \bar{H}_k / \log K$; higher = better separation).

```{r supp-table3}
post <- as.data.frame(mod_train$z)
colnames(post) <- paste0("p_cl", seq_len(ncol(post)))

K <- mod_train$G
dat_post <- mydata %>%
  bind_cols(post) %>%
  mutate(
    max_prob = apply(select(., starts_with("p_cl")), 1, max),
    entropy  = apply(select(., starts_with("p_cl")), 1,
                     function(p) -sum(ifelse(p > 0, p * log(p), 0)))
  )

supp_table3 <- dat_post %>%
  group_by(cluster_assignment) %>%
  summarise(
    n                    = n(),
    mean_max_posterior   = mean(max_prob),
    median_max_posterior = median(max_prob),
    mean_entropy         = mean(entropy),
    relative_entropy     = 1 - mean(entropy) / log(K),
    .groups = "drop"
  )

print(supp_table3, digits = 3)
write.csv(supp_table3,
          file.path(TABLES_DIR, "supp_table3_lpa_classification.csv"),
          row.names = FALSE)
```

# Overall ATE checks

```{r overall-ate}
# 1) Any SSI vs control
lm_binary <- lm(phq9_score_followup ~ phq9_pre_rev + int_binary,
                data = mydata, weights = w_norm)
cat("=== Any SSI vs Control ===\n")
print(tidy(lm_binary) %>% filter(term == "int_binary"), digits = 3)

# 2) 12-arm omnibus among intervention arms only
dat_int <- subset(mydata, group_numeric != "13")
dat_int$group_int_f <- factor(dat_int$group_numeric)
lm_int_only <- lm(phq9_score_followup ~ phq9_pre_rev + group_int_f,
                  data = dat_int, weights = w_norm)
cat("\n=== Omnibus across 12 SSIs (intervention arms only) ===\n")
print(car::Anova(lm_int_only, type = 2))
```

## Supplementary Table 4 --- Per-arm vs control with FDR

```{r supp-table4}
mydata$group_f <- factor(mydata$group_numeric)
mydata$group_f <- relevel(mydata$group_f, ref = "13")

lm_main   <- lm(phq9_score_followup ~ phq9_pre_rev + group_f,
                data = mydata, weights = w_norm)
coef_tab  <- summary(lm_main)$coefficients
ssi_rows  <- grep("^group_f", rownames(coef_tab))

supp_table4 <- data.frame(
  ssi    = gsub("group_f", "SSI ", rownames(coef_tab)[ssi_rows]),
  b      = coef_tab[ssi_rows, "Estimate"],
  se     = coef_tab[ssi_rows, "Std. Error"],
  t      = coef_tab[ssi_rows, "t value"],
  p_raw  = coef_tab[ssi_rows, "Pr(>|t|)"]
) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH")) %>%
  arrange(p_raw) %>%
  mutate(
    ssi_num = as.integer(gsub("SSI ", "", ssi)),
    ssi     = paste0(ssi_num, ". ", ssi_labels[ssi_num])
  ) %>%
  select(-ssi_num)

print(supp_table4, digits = 3)
write.csv(supp_table4,
          file.path(TABLES_DIR, "supp_table4_per_arm_fdr.csv"),
          row.names = FALSE)
```

# Cluster-specific intervention effects → Table 3

```{r table3}
targets <- list(
  "1" = c("1", "5", "7", "8", "9"),
  "2" = c("2", "3", "6", "9", "11"),
  "3" = c("10", "12"),
  "4" = c("3", "5")
)

mydata$cluster_assignment <- factor(mydata$cluster_assignment)
mydata$group_numeric_chr  <- as.character(mydata$group_numeric)

cluster_subsets <- list(
  d1 = subset(mydata, cluster_assignment == "1"),
  d2 = subset(mydata, cluster_assignment == "2"),
  d3 = subset(mydata, cluster_assignment == "3"),
  d4 = subset(mydata, cluster_assignment == "4")
)

# Helper closures
fit_binary <- function(d) {
  d$int_binary <- ifelse(d$group_numeric_chr == "13", 0, 1)
  mod <- lm(phq9_score_followup ~ phq9_pre_rev + int_binary,
            data = d, weights = w_norm)
  tidy(mod) %>% filter(term == "int_binary")
}

fit_target <- function(d, tgt) {
  d$int_binary <- ifelse(d$group_numeric_chr == "13", 0, 1)
  d$int_target <- ifelse(d$group_numeric_chr %in% tgt, 1, 0)
  mod <- lm(phq9_score_followup ~ phq9_pre_rev + int_binary + int_target,
            data = d, weights = w_norm)
  tidy(mod) %>% filter(term == "int_target")
}

fit_emm_pairs <- function(d, tgt) {
  d$trt_3cat <- factor(
    ifelse(d$group_numeric_chr == "13", "control",
           ifelse(d$group_numeric_chr %in% tgt, "target", "other_int")),
    levels = c("control", "other_int", "target")
  )
  mod <- lm(phq9_score_followup ~ phq9_pre_rev + trt_3cat,
            data = d, weights = w_norm)
  # reverse = TRUE: contrasts read intervention - control (positive = improvement)
  pairs(emmeans(mod, specs = "trt_3cat"), reverse = TRUE) %>% as.data.frame()
}

fit_ftest <- function(d, tgt) {
  d$int_binary <- ifelse(d$group_numeric_chr == "13", 0, 1)
  d$int_target <- ifelse(d$group_numeric_chr %in% tgt, 1, 0)
  mod_base   <- lm(phq9_score_followup ~ phq9_pre_rev + int_binary,
                   data = d, weights = w_norm)
  mod_target <- lm(phq9_score_followup ~ phq9_pre_rev + int_binary + int_target,
                   data = d, weights = w_norm)
  a <- anova(mod_base, mod_target)
  data.frame(delta_F = a$F[2], df1 = a$Df[2], df_resid = a$Res.Df[2],
             p.value = a$`Pr(>F)`[2])
}

# Format helpers
fmt_p_tbl <- function(p) {
  if (is.na(p)) return(NA_character_)
  if (p < .001) return("< .001")
  sub("^0\\.", ".", sprintf("%.3f", p))
}
n_completers <- function(d) sum(!is.na(d$phq9_score_followup))
indent <- strrep("\u00A0", 4)   # non-breaking spaces so sub-rows indent in HTML/Word

# Compute models for all clusters
m1 <- lapply(seq_along(cluster_subsets),
             function(i) fit_binary(cluster_subsets[[i]]))
m2 <- lapply(seq_along(cluster_subsets),
             function(i) fit_target(cluster_subsets[[i]],
                                    targets[[as.character(i)]]))
emm_list <- lapply(seq_along(cluster_subsets),
                   function(i) fit_emm_pairs(cluster_subsets[[i]],
                                             targets[[as.character(i)]]))
ftests <- lapply(seq_along(cluster_subsets),
                 function(i) fit_ftest(cluster_subsets[[i]],
                                       targets[[as.character(i)]]))

# Cluster metadata
cluster_meta <- list(
  list(label = "Cluster 1 (primary)",     targets_str = "1, 5, 7, 8, 9"),
  list(label = "Cluster 2 (exploratory)", targets_str = "2, 3, 6, 9, 11"),
  list(label = "Cluster 3 (primary)",     targets_str = "10, 12"),
  list(label = "Cluster 4 (exploratory)", targets_str = "3, 5")
)

# Build long-format table (Metric | b | SE | t | p)
long_rows <- list()
for (i in 1:4) {
  n_c <- n_completers(cluster_subsets[[i]])
  ft  <- ftests[[i]]
  hdr <- sprintf("%s (n = %d; target SSIs: %s)",
                 cluster_meta[[i]]$label, n_c, cluster_meta[[i]]$targets_str)

  long_rows[[length(long_rows) + 1]] <- data.frame(
    Metric = hdr, b = NA_real_, SE = NA_real_, t = NA_real_,
    p = NA_character_, stringsAsFactors = FALSE)

  long_rows[[length(long_rows) + 1]] <- data.frame(
    Metric = paste0(indent, "Any SSI vs. control"),
    b  = round(m1[[i]]$estimate,  2),
    SE = round(m1[[i]]$std.error, 2),
    t  = round(m1[[i]]$statistic, 2),
    p  = fmt_p_tbl(m1[[i]]$p.value),
    stringsAsFactors = FALSE)

  long_rows[[length(long_rows) + 1]] <- data.frame(
    Metric = paste0(indent, "Target SSI (incremental)"),
    b  = round(m2[[i]]$estimate,  2),
    SE = round(m2[[i]]$std.error, 2),
    t  = round(m2[[i]]$statistic, 2),
    p  = fmt_p_tbl(m2[[i]]$p.value),
    stringsAsFactors = FALSE)

  long_rows[[length(long_rows) + 1]] <- data.frame(
    Metric = sprintf("%sModel fit improvement: ΔF(%d, %d) = %.2f",
                     indent, ft$df1, ft$df_resid, ft$delta_F),
    b = NA_real_, SE = NA_real_, t = NA_real_,
    p = fmt_p_tbl(ft$p.value),
    stringsAsFactors = FALSE)

  for (cname in c("target - control", "other_int - control", "target - other_int")) {
    lbl <- switch(cname,
      "target - control"    = paste0(indent, "Target SSIs vs. control"),
      "other_int - control" = paste0(indent, "Other SSIs vs. control"),
      "target - other_int"  = paste0(indent, "Target SSIs vs. other SSIs"))
    crow <- emm_list[[i]] %>% filter(contrast == cname)
    if (nrow(crow) == 0) next
    long_rows[[length(long_rows) + 1]] <- data.frame(
      Metric = lbl,
      b  = round(crow$estimate, 2),
      SE = round(crow$SE,       2),
      t  = round(crow$t.ratio,  2),
      p  = fmt_p_tbl(crow$p.value),
      stringsAsFactors = FALSE)
  }
}

table3 <- do.call(rbind, long_rows)
rownames(table3) <- NULL
print(table3)

write.csv(table3,
          file.path(TABLES_DIR, "table3_cluster_effects.csv"),
          row.names = FALSE)

# Cluster × target-intervention interaction (Cluster 1 follow-up reported in text)
dat_target <- subset(mydata, group_numeric_chr == "13" |
                              group_numeric_chr %in% targets[["1"]])
dat_target$int_target <- ifelse(dat_target$group_numeric_chr == "13", 0, 1)
lm_target_cluster <- lm(
  phq9_score_followup ~ phq9_pre_rev + cluster_assignment * int_target,
  data    = dat_target,
  weights = w_norm
)
cat("\n=== Cluster × Target-intervention (Cluster 1 targets) interaction ===\n")
print(car::Anova(lm_target_cluster, type = 2))
print(pairs(emmeans(lm_target_cluster, ~ int_target | cluster_assignment), reverse = TRUE))

# Cluster × target-intervention interaction (Cluster 3 follow-up reported in text)
dat_target_c3 <- subset(mydata, group_numeric_chr == "13" |
                                 group_numeric_chr %in% targets[["3"]])
dat_target_c3$int_target <- ifelse(dat_target_c3$group_numeric_chr == "13", 0, 1)
lm_target_cluster_c3 <- lm(
  phq9_score_followup ~ phq9_pre_rev + cluster_assignment * int_target,
  data    = dat_target_c3,
  weights = w_norm
)
cat("\n=== Cluster × Target-intervention (Cluster 3 targets) interaction ===\n")
print(car::Anova(lm_target_cluster_c3, type = 2))
print(pairs(emmeans(lm_target_cluster_c3, ~ int_target | cluster_assignment), reverse = TRUE))
```

# Cluster characterization --- LASSO

Binary LASSO logistic regressions (cluster *k* vs. rest) with 10-fold CV (`tidymodels`/`glmnet`), using baseline scale totals plus demographics; item-level covariates that roll up into totals are dropped.

```{r lasso-setup}
dat_post <- dat_post %>%
  mutate(
    hopeless_score_baseline = q1_beck_baseline + q2_beck_baseline +
                              q3_beck_baseline + q4_beck_baseline,
    fats_score_baseline     = rowSums(across(fats_1_baseline:fats_12_baseline)),
    readiness_score_baseline = ready_important_baseline + ready_confident_baseline,
    agency_score_baseline    = agency_q1_baseline + agency_q2_baseline +
                               agency_q3_baseline,
    dep_expectancies_score_baseline = expect_donthelp_r_baseline +
                                       expect_wontwork_r_baseline +
                                       expect_positive_baseline
  )

drop_items <- c(
  paste0("phq9_", 1:9, "_baseline"),
  paste0("q", 1:4, "_beck_baseline"),
  paste0("fats_", 1:12, "_baseline"),
  paste0("agency_q", 1:3, "_baseline"),
  c("ready_important_baseline", "ready_confident_baseline",
    "expect_donthelp_r_baseline", "expect_wontwork_r_baseline",
    "expect_positive_baseline")
)

scale_totals <- c(
  "phq9_pre_rev", "hopeless_score_baseline", "fats_score_baseline",
  "readiness_score_baseline", "agency_score_baseline",
  "dep_expectancies_score_baseline", "ffmq_distressing_baseline"
)

lasso_predictors <- function(data) {
  # Defense-in-depth: also strip any `c\d+_binary` cluster indicators that
  # might have been added downstream — they would leak the outcome.
  drop_cols <- c("pid", "w_norm", "cluster_assignment",
                 "phq9_score_followup", "fup_complete", "int_binary",
                 "phq9_pre_raw", "group_numeric", "group_f",
                 "group_numeric_chr",
                 grep("^p_cl",       names(data), value = TRUE),
                 grep("^c\\d+_binary$", names(data), value = TRUE),
                 "max_prob", "entropy")
  candidates <- setdiff(names(data),
                        c(drop_cols, intersect(names(data), drop_items)))
  unique(c(candidates, intersect(scale_totals, names(data))))
}

run_lasso <- function(data, target_cluster, seed = 123) {
  preds <- lasso_predictors(data)
  d <- data %>%
    mutate(target = factor(if_else(cluster_assignment == target_cluster,
                                    "Target", "Other"))) %>%
    select(target, all_of(preds))

  set.seed(seed)
  rec <- recipe(target ~ ., data = d) %>%
    step_nzv(all_predictors()) %>%
    step_dummy(all_nominal_predictors()) %>%
    step_normalize(all_numeric_predictors())
  spec <- logistic_reg(mode = "classification",
                       penalty = tune(), mixture = 1) %>%
    set_engine("glmnet")
  wf    <- workflow() %>% add_recipe(rec) %>% add_model(spec)
  folds <- vfold_cv(d, v = 10, strata = target)
  grid  <- grid_regular(penalty(), levels = 60)
  res   <- tune_grid(wf, resamples = folds, grid = grid,
                     metrics = metric_set(roc_auc))
  best  <- select_best(res, metric = "roc_auc")
  fit_final <- finalize_workflow(wf, best) %>% fit(d)
  tidy(fit_final) %>%
    filter(term != "(Intercept)", estimate != 0) %>%
    mutate(abs_est = abs(estimate)) %>%
    arrange(desc(abs_est))
}
```

## Cluster 1 LASSO

```{r lasso-c1}
var_labels <- c(
  phq9_pre_rev                    = "Baseline PHQ-9 (reverse scored)",
  fats_score_baseline             = "FATS",
  agency_score_baseline           = "Agency — State Hope Scale",
  hopeless_score_baseline         = "Hopelessness — BHS-4",
  dep_expectancies_score_baseline = "Depression change expectancies — DCES",
  age_baseline                    = "Age",
  income_ordered_baseline         = "Household income (ordered)",
  social_ladder_baseline          = "MacArthur Social Ladder",
  education_ordered_baseline      = "Education (ordered)",
  disability_yes_baseline         = "Disability or chronic condition",
  help_mental_health_app_baseline = "Prior mental health app use",
  help_online_resource_baseline   = "Prior online mental health resource use",
  gender_woman_baseline           = "Gender: woman",
  gender_man_baseline             = "Gender: man",
  political_party_independent_baseline    = "Political Party: Independent",
  political_party_republican_baseline     = "Political Party: Republican",
  political_party_something_else_baseline = "Political Party: Something else",
  race_white_baseline                     = "Race: White",
  race_black_baseline                     = "Race: Black",
  ethnicity_not_hispanic_baseline         = "Ethnicity: Not Hispanic/Latino",
  ethnicity_mexican_baseline              = "Ethnicity: Mexican",
  employment_student_baseline             = "Employment: Student",
  employment_full_time_baseline           = "Employment: Full-time",
  employment_part_time_baseline           = "Employment: Part-time",
  employment_unemployed_baseline          = "Employment: Unemployed",
  employment_not_in_paid_work_baseline    = "Employment: Not in paid work",
  marital_status_in_relationship_baseline = "Marital Status: In a relationship",
  marital_status_married_baseline         = "Marital Status: Married",
  marital_status_divorced_baseline        = "Marital Status: Divorced",
  help_support_group_baseline             = "Prior help: Support group",
  help_phone_text_baseline                = "Prior help: Phone/text line",
  help_therapist_baseline                 = "Prior help: Therapist",
  help_medical_doctor_baseline            = "Prior help: Medical doctor",
  help_book_video_baseline                = "Prior help: Book/video",
  help_no_formal_support_baseline         = "Prior help: No formal support",
  readiness_score_baseline                = "Readiness to change",
  ffmq_distressing_baseline               = "FFMQ — Non-reactivity to distressing thoughts",
  recruitment_platform_prolific_baseline  = "Recruitment platform: Prolific"
)
relabel_terms <- function(df) {
  df %>% mutate(term = ifelse(
    term %in% names(var_labels),
    var_labels[term],
    gsub("_baseline$", "", gsub("_", " ", term))
  ))
}

lasso_c1 <- run_lasso(dat_post, target_cluster = 1, seed = 123)
print(lasso_c1, n = Inf)
write.csv(relabel_terms(lasso_c1),
          file.path(TABLES_DIR, "supp_table6_lasso_cluster1.csv"),
          row.names = FALSE)
```

## Cluster 3 LASSO

```{r lasso-c3}
lasso_c3 <- run_lasso(dat_post, target_cluster = 3, seed = 123)
print(lasso_c3, n = Inf)
write.csv(relabel_terms(lasso_c3),
          file.path(TABLES_DIR, "supp_table7_lasso_cluster3.csv"),
          row.names = FALSE)
```

# SSI characteristics by target / non-target → Supplementary Table 8

```{r ssi-characteristics}
ssi_char <- read.csv(file.path(DATA_DIR, "ssi_characteristics.csv"),
                     header = TRUE) %>%
  mutate(group_numeric = case_when(
    str_detect(Program_Name, "5 Habits")              ~ 1,
    str_detect(Program_Name, "AI-assisted")           ~ 2,
    str_detect(Program_Name, "Behavioral Activation") ~ 3,
    str_detect(Program_Name, "Dialectical")           ~ 4,
    str_detect(Program_Name, "Inner Child")           ~ 5,
    str_detect(Program_Name, "Interactive Cognitive") ~ 6,
    str_detect(Program_Name, "Mindful Acceptance")    ~ 7,
    str_detect(Program_Name, "Mindful Attention")     ~ 8,
    str_detect(Program_Name, "Moral Elevation")       ~ 9,
    str_detect(Program_Name, "Personalized Intervention") ~ 10,
    str_detect(Program_Name, "Reframing Negative")    ~ 11,
    str_detect(Program_Name, "Savoring")              ~ 12,
    TRUE ~ NA_real_
  ))

continuous_vars <- c("Engagingness", "Interactive_Activities", "Effortful",
                     "Personalization", "Cognitive_Reframing",
                     "Behavioral_Activation", "Mindfulness")
binary_vars     <- c("Video_Audio", "References_Depression", "Links_to_Resources",
                     "Meditation_Practice", "Clinician_Created", "Take_Home_Reminder")

df_char <- ssi_char %>%
  mutate(
    target_c1 = if_else(group_numeric %in% c(1, 5, 7, 8, 9), "Target", "Other"),
    target_c3 = if_else(group_numeric %in% c(10, 12),       "Target", "Other"),
    across(all_of(binary_vars), ~ if_else(. == "Yes", 1L, 0L))
  )

summarize_by_target <- function(data, target_var) {
  cont_tbl <- data %>%
    group_by({{ target_var }}) %>%
    summarise(across(all_of(continuous_vars),
                     ~ sprintf("%.2f (%.2f)", mean(., na.rm = TRUE),
                                              sd(., na.rm = TRUE))),
              .groups = "drop") %>%
    pivot_longer(-{{ target_var }}, names_to = "Characteristic", values_to = "val") %>%
    pivot_wider(names_from = {{ target_var }}, values_from = "val") %>%
    mutate(type = "continuous")

  bin_tbl <- data %>%
    group_by({{ target_var }}) %>%
    summarise(across(all_of(binary_vars),
                     ~ sprintf("%d (%.0f%%)", sum(., na.rm = TRUE),
                                              100 * mean(., na.rm = TRUE))),
              .groups = "drop") %>%
    pivot_longer(-{{ target_var }}, names_to = "Characteristic", values_to = "val") %>%
    pivot_wider(names_from = {{ target_var }}, values_from = "val") %>%
    mutate(type = "binary")

  bind_rows(cont_tbl, bin_tbl) %>%
    arrange(type, Characteristic)
}

table_c1 <- summarize_by_target(df_char, target_c1) %>%
  rename(Other_C1 = Other, Target_C1 = Target)
table_c3 <- summarize_by_target(df_char, target_c3) %>%
  rename(Other_C3 = Other, Target_C3 = Target)

supp_table7 <- table_c1 %>%
  full_join(table_c3, by = c("Characteristic", "type"))

char_labels <- c(
  Engagingness           = "Engagingness",
  Interactive_Activities = "Interactive Activities",
  Effortful              = "Effortful",
  Personalization        = "Personalization",
  Cognitive_Reframing    = "Cognitive Reframing",
  Behavioral_Activation  = "Behavioral Activation",
  Mindfulness            = "Mindfulness",
  Video_Audio            = "Video/Audio",
  References_Depression  = "References Depression",
  Links_to_Resources     = "Links to Resources",
  Meditation_Practice    = "Meditation Practice",
  Clinician_Created      = "Clinician Created",
  Take_Home_Reminder     = "Take-Home Reminder"
)
supp_table7 <- supp_table7 %>%
  mutate(Characteristic = ifelse(Characteristic %in% names(char_labels),
                                  char_labels[Characteristic],
                                  Characteristic)) %>%
  select(-type)

print(supp_table7)
write.csv(supp_table7,
          file.path(TABLES_DIR, "supp_table8_ssi_characteristics_by_target.csv"),
          row.names = FALSE)
```

# Assignment-strategy comparison

Each strategy is benchmarked against control via a weighted regression of reverse-scored 4-week PHQ-9 on baseline PHQ-9 and a binary treatment indicator. HTE-informed strategies multiply the IPW weight by a design weight that corrects for the number of selected arms per cluster (`12 / k`) and re-calibrates the filtered sample's cluster × treatment proportions to the full completer sample.

```{r strategy-helpers}
# Step 1: full-sample target proportions
full_wtd_props <- mydata %>%
  filter(!is.na(phq9_score_followup)) %>%
  mutate(int_binary = ifelse(group_numeric_chr == "13", 0, 1)) %>%
  group_by(cluster_assignment, int_binary) %>%
  summarise(wsum_full = sum(w_norm), .groups = "drop") %>%
  mutate(wprop_full = wsum_full / sum(wsum_full))

# Single-strategy fitter that handles arm-count + ratio calibration
fit_strategy <- function(d, n_arms_lookup) {
  d <- d %>%
    filter(!is.na(phq9_score_followup)) %>%
    mutate(int_opt = ifelse(group_numeric_chr == "13", 0, 1)) %>%
    mutate(arm_correction = ifelse(int_opt == 1,
                                   12 / n_arms_lookup[as.character(cluster_assignment)],
                                   1),
           w_arm = w_norm * arm_correction)

  filt_props <- d %>%
    group_by(cluster_assignment, int_opt) %>%
    summarise(wsum_filt = sum(w_arm), .groups = "drop") %>%
    mutate(wprop_filt = wsum_filt / sum(wsum_filt)) %>%
    rename(int_binary = int_opt)

  d <- d %>%
    left_join(full_wtd_props %>% select(cluster_assignment, int_binary, wprop_full),
              by = c("cluster_assignment", "int_binary" = "int_binary")) %>%
    left_join(filt_props %>% select(cluster_assignment, int_binary, wprop_filt),
              by = c("cluster_assignment", "int_binary" = "int_binary")) %>%
    mutate(ratio = wprop_full / wprop_filt,
           w_combined = w_arm * ratio)

  mod <- lm(phq9_score_followup ~ phq9_pre_rev + int_opt,
            data = d, weights = w_combined)
  list(d = d, mod = mod,
       row = tidy(mod) %>% filter(term == "int_opt"))
}
```

```{r strategy-fits}
# Strategy 1: Random — any SSI vs control across the full sample
mydata_complete <- mydata %>% filter(!is.na(phq9_score_followup))
lm_s1 <- lm(phq9_score_followup ~ phq9_pre_rev + int_binary,
            data = mydata_complete, weights = w_norm)
s1_row <- tidy(lm_s1) %>% filter(term == "int_binary") %>%
  mutate(strategy = "1_random_assignment")

# Strategy 2: Best aggregate — SSIs 6 & 8 vs control
d_s2 <- mydata_complete %>%
  filter(group_numeric_chr %in% c("6", "8", "13")) %>%
  mutate(int_agg = ifelse(group_numeric_chr == "13", 0, 1))
lm_s2 <- lm(phq9_score_followup ~ phq9_pre_rev + int_agg,
            data = d_s2, weights = w_norm)
s2_row <- tidy(lm_s2) %>% filter(term == "int_agg") %>%
  mutate(strategy = "2_best_aggregate")

# Strategy 3: HTE-optimal in C1/C3, aggregate (6,8) in C2/C4
d_s3 <- mydata %>%
  filter(
    (cluster_assignment == "1" & group_numeric_chr %in% c("1","5","7","8","9","13")) |
    (cluster_assignment == "2" & group_numeric_chr %in% c("6","8","13"))             |
    (cluster_assignment == "3" & group_numeric_chr %in% c("10","12","13"))           |
    (cluster_assignment == "4" & group_numeric_chr %in% c("6","8","13"))
  )
fit_s3 <- fit_strategy(d_s3, c("1" = 5, "2" = 2, "3" = 2, "4" = 2))
s3_row <- fit_s3$row %>% mutate(strategy = "3_hte_plus_aggregate")

strategy_table <- bind_rows(s1_row, s2_row, s3_row) %>%
  transmute(strategy,
            b      = round(estimate,  3),
            se     = round(std.error, 3),
            t      = round(statistic, 3),
            p      = round(p.value,   4))

print(strategy_table)
```

## Bootstrap pairwise contrasts (5,000 resamples)

```{r bootstrap}
# Each bootstrap statistic re-fits both strategies on the resampled data
# and returns the difference in coefficients.

boot_diff_1v2 <- function(data, indices) {
  d <- data[indices, ]
  b1 <- coef(lm(phq9_score_followup ~ phq9_pre_rev + int_binary,
                data = d, weights = w_norm))["int_binary"]
  d_agg <- d %>% filter(group_numeric_chr %in% c("6","8","13")) %>%
    mutate(int_agg = ifelse(group_numeric_chr == "13", 0, 1))
  b2 <- coef(lm(phq9_score_followup ~ phq9_pre_rev + int_agg,
                data = d_agg, weights = w_norm))["int_agg"]
  unname(b2 - b1)
}

boot_strategy_3 <- function(d) {
  n_arms <- c("1" = 5, "2" = 2, "3" = 2, "4" = 2)
  d_opt <- d %>%
    filter(
      (cluster_assignment == "1" & group_numeric_chr %in% c("1","5","7","8","9","13")) |
      (cluster_assignment == "2" & group_numeric_chr %in% c("6","8","13"))             |
      (cluster_assignment == "3" & group_numeric_chr %in% c("10","12","13"))           |
      (cluster_assignment == "4" & group_numeric_chr %in% c("6","8","13"))
    ) %>%
    mutate(int_opt = ifelse(group_numeric_chr == "13", 0, 1),
           arm_correction = ifelse(int_opt == 1,
                                   12 / n_arms[as.character(cluster_assignment)],
                                   1),
           w_arm = w_norm * arm_correction)
  filt_props <- d_opt %>%
    group_by(cluster_assignment, int_opt) %>%
    summarise(wsum_filt = sum(w_arm), .groups = "drop") %>%
    mutate(wprop_filt = wsum_filt / sum(wsum_filt)) %>%
    rename(int_binary = int_opt)
  d_opt <- d_opt %>%
    left_join(full_wtd_props %>% select(cluster_assignment, int_binary, wprop_full),
              by = c("cluster_assignment", "int_binary")) %>%
    left_join(filt_props %>% select(cluster_assignment, int_binary, wprop_filt),
              by = c("cluster_assignment", "int_binary")) %>%
    mutate(ratio = wprop_full / wprop_filt,
           w_combined = w_arm * ratio)
  unname(coef(lm(phq9_score_followup ~ phq9_pre_rev + int_opt,
                 data = d_opt, weights = w_combined))["int_opt"])
}

boot_diff_1v3 <- function(data, indices) {
  d <- data[indices, ]
  b1 <- coef(lm(phq9_score_followup ~ phq9_pre_rev + int_binary,
                data = d, weights = w_norm))["int_binary"]
  boot_strategy_3(d) - unname(b1)
}

boot_diff_2v3 <- function(data, indices) {
  d <- data[indices, ]
  d_agg <- d %>% filter(group_numeric_chr %in% c("6","8","13")) %>%
    mutate(int_agg = ifelse(group_numeric_chr == "13", 0, 1))
  b2 <- coef(lm(phq9_score_followup ~ phq9_pre_rev + int_agg,
                data = d_agg, weights = w_norm))["int_agg"]
  boot_strategy_3(d) - unname(b2)
}

bootstrap_results <- {
  set.seed(12345)
  list(
    `1v2` = boot(mydata_complete, statistic = boot_diff_1v2, R = 5000),
    `1v3` = boot(mydata_complete, statistic = boot_diff_1v3, R = 5000),
    `2v3` = boot(mydata_complete, statistic = boot_diff_2v3, R = 5000)
  )
}

summarize_boot <- function(b, label) {
  ci <- boot.ci(b, type = "perc")
  z  <- b$t0 / sd(b$t)
  data.frame(
    contrast      = label,
    diff          = round(b$t0,            3),
    boot_se       = round(sd(b$t),         3),
    ci_lo         = round(ci$percent[4],   3),
    ci_hi         = round(ci$percent[5],   3),
    p_two_sided   = round(2 * pnorm(-abs(z)), 4)
  )
}

bootstrap_table <- bind_rows(
  summarize_boot(bootstrap_results[["1v2"]], "Strategy 2 - Strategy 1"),
  summarize_boot(bootstrap_results[["1v3"]], "Strategy 3 - Strategy 1"),
  summarize_boot(bootstrap_results[["2v3"]], "Strategy 3 - Strategy 2")
)
print(bootstrap_table)
```

## Combined Supplementary Table 8 (strategies + bootstrap)

```{r supp-table8}
strategy_block <- strategy_table %>%
  transmute(row_type = "strategy_estimate",
            label    = strategy,
            estimate = b,
            se,
            ci_lo    = NA_real_,
            ci_hi    = NA_real_,
            stat     = t,
            p)

bootstrap_block <- bootstrap_table %>%
  transmute(row_type = "bootstrap_contrast",
            label    = contrast,
            estimate = diff,
            se       = boot_se,
            ci_lo,
            ci_hi,
            stat     = NA_real_,
            p        = p_two_sided)

supp_table8 <- bind_rows(strategy_block, bootstrap_block)
print(supp_table8)
write.csv(supp_table8,
          file.path(TABLES_DIR, "supp_table8_strategy_comparison.csv"),
          row.names = FALSE)
```

# Demographic Table 1 (descriptive)

Baseline descriptives (N = 7,505), weighted follow-up, and weighted within-arm range.

```{r table1}
# ---- Derived vars: 4-cat race/ethnicity + 4-cat education + analysis scales ----
t1 <- mydata %>%
  mutate(
    phq9_baseline = phq9_pre_raw,
    fats_score    = rowSums(across(fats_1_baseline:fats_12_baseline), na.rm = TRUE),
    agency_score  = agency_q1_baseline + agency_q2_baseline + agency_q3_baseline,
    bhs4_score    = q1_beck_baseline + q2_beck_baseline +
                     q3_beck_baseline + q4_beck_baseline,
    dces_score    = expect_donthelp_r_baseline + expect_wontwork_r_baseline +
                     expect_positive_baseline,
    gender_nonbinary_other = as.integer(rowSums(across(c(
      gender_non_binary_baseline, gender_additional_gender_baseline,
      gender_agender_baseline, gender_two_spirit_baseline)), na.rm = TRUE) > 0),
    employment_other = as.integer(rowSums(across(c(
      employment_business_owner_baseline, employment_retired_baseline,
      employment_prefer_not_say_baseline)), na.rm = TRUE) > 0),

    # Race is a single-select survey item, so categories are mutually
    # exclusive and exhaustive.
    race_asian = as.integer(rowSums(across(c(
      race_asian_indian_baseline, race_chinese_baseline, race_filipino_baseline,
      race_vietnamese_baseline, race_korean_baseline, race_japanese_baseline,
      race_asian_other_baseline)), na.rm = TRUE) > 0),
    race_napi = as.integer(rowSums(across(c(
      race_native_american_baseline, race_hawaiian_baseline,
      race_guamanian_baseline, race_samoan_baseline)), na.rm = TRUE) > 0),
    # Hispanic or Latino ethnicity (separate survey item from race; yes/no)
    hispanic = as.integer(rowSums(across(c(
      ethnicity_mexican_baseline, ethnicity_puerto_rican_baseline,
      ethnicity_cuban_baseline, ethnicity_other_hispanic_baseline)),
      na.rm = TRUE) > 0),

    # 4-category education from the 0-8 ordered scale
    educ_lt_hs   = as.integer(education_ordered_baseline %in% c(0, 1)),
    educ_hs_ged  = as.integer(education_ordered_baseline == 2),
    educ_some_ps = as.integer(education_ordered_baseline %in% c(3, 4)),
    educ_college = as.integer(education_ordered_baseline %in% c(5, 6, 7, 8)),

    # 4-band household income from the 0-16 ordered bracket scale
    inc_lt40    = as.integer(income_ordered_baseline %in% 0:3),
    inc_40_80   = as.integer(income_ordered_baseline %in% 4:7),
    inc_80_100  = as.integer(income_ordered_baseline %in% 8:9),
    inc_100plus = as.integer(income_ordered_baseline %in% 10:16),
    inc_prefer  = as.integer(is.na(income_ordered_baseline))
  )

comp      <- t1 %>% filter(fup_complete == 1)             # weighted FU sample
arm_order <- c(as.character(1:12), "13")
arm_names <- c(ssi_labels, "Passive control")

# ---- weighted helpers + one-decimal cell formatters ----
wmean <- function(x, w) { ok <- !is.na(x) & !is.na(w); sum(x[ok]*w[ok])/sum(w[ok]) }
wsd   <- function(x, w) { ok <- !is.na(x) & !is.na(w); m <- wmean(x, w)
                          sqrt(sum(w[ok]*(x[ok]-m)^2)/sum(w[ok])) }
cell_cont <- function(x, w = NULL) if (is.null(w))
  sprintf("%.1f (%.1f)", mean(x, na.rm = TRUE), sd(x, na.rm = TRUE)) else
  sprintf("%.1f (%.1f)", wmean(x, w), wsd(x, w))
cell_bin  <- function(x, w = NULL) if (is.null(w))
  sprintf("%.1f", mean(x == 1, na.rm = TRUE) * 100) else
  sprintf("%.1f", wmean(as.integer(x == 1), w) * 100)
est_num   <- function(type, x, w) if (type == "cont") wmean(x, w) else
  wmean(as.integer(x == 1), w) * 100
fmt_row   <- function(type, col, d, w) {
  if (type == "header") return("")
  if (type == "cont") cell_cont(d[[col]], w) else cell_bin(d[[col]], w)
}

# ---- row spec: display label, type, source column ----
ind <- strrep("\u00A0", 4)   # non-breaking spaces so sub-rows indent in docx
items <- tibble::tribble(
  ~label,                                            ~type,    ~col,
  "Age, M (SD)",                                     "cont",   "age_baseline",
  "Gender, %",                                       "header", NA,
  paste0(ind, "Woman"),                              "bin",    "gender_woman_baseline",
  paste0(ind, "Man"),                                "bin",    "gender_man_baseline",
  paste0(ind, "Non-binary or other gender identity"), "bin",  "gender_nonbinary_other",
  "Race, %",                                         "header", NA,
  paste0(ind, "White"),                              "bin",    "race_white_baseline",
  paste0(ind, "Black or African American"),          "bin",    "race_black_baseline",
  paste0(ind, "Asian"),                              "bin",    "race_asian",
  paste0(ind, "American Indian, Alaska Native, or Pacific Islander"), "bin", "race_napi",
  paste0(ind, "Other"),                              "bin",    "race_other_baseline",
  paste0(ind, "Prefer not to say"),                  "bin",    "race_prefer_not_say_baseline",
  "Hispanic or Latino, %",                           "bin",    "hispanic",
  "Education, %",                                     "header", NA,
  paste0(ind, "Less than high school"),              "bin",    "educ_lt_hs",
  paste0(ind, "High school graduate or GED"),        "bin",    "educ_hs_ged",
  paste0(ind, "Some post-secondary"),                "bin",    "educ_some_ps",
  paste0(ind, "College graduate"),                   "bin",    "educ_college",
  "Household income (USD), %",                        "header", NA,
  paste0(ind, "Less than 40,000"),                   "bin",    "inc_lt40",
  paste0(ind, "40,000 to 79,999"),                   "bin",    "inc_40_80",
  paste0(ind, "80,000 to 99,999"),                   "bin",    "inc_80_100",
  paste0(ind, "100,000 or more"),                    "bin",    "inc_100plus",
  paste0(ind, "Prefer not to say"),                  "bin",    "inc_prefer",
  "MacArthur Social Ladder (1-10), M (SD)",          "cont",   "social_ladder_baseline",
  "Employment status, %",                            "header", NA,
  paste0(ind, "Full-time employed"),                 "bin",    "employment_full_time_baseline",
  paste0(ind, "Part-time employed"),                 "bin",    "employment_part_time_baseline",
  paste0(ind, "Unemployed"),                         "bin",    "employment_unemployed_baseline",
  paste0(ind, "Not in paid work"),                   "bin",    "employment_not_in_paid_work_baseline",
  paste0(ind, "Student"),                            "bin",    "employment_student_baseline",
  paste0(ind, "Other or prefer not to say"),         "bin",    "employment_other",
  "Disability or chronic condition, %",              "bin",    "disability_yes_baseline",
  "Prior mental health app use, %",                  "bin",    "help_mental_health_app_baseline",
  "Prior online mental health resource use, %",      "bin",    "help_online_resource_baseline",
  "Baseline PHQ-9, M (SD)",                          "cont",   "phq9_baseline",
  "FATS, M (SD)",                                     "cont",   "fats_score",
  "Agency (State Hope Scale), M (SD)",               "cont",   "agency_score",
  "Hopelessness (BHS-4), M (SD)",                    "cont",   "bhs4_score",
  "Depression change expectancies (DCES), M (SD)",   "cont",   "dces_score"
)

# ---- compute the three summary columns ----
build <- items %>%
  mutate(
    baseline = vapply(seq_len(n()), function(i)
      fmt_row(type[i], col[i], t1, NULL), character(1)),
    weighted_fu = vapply(seq_len(n()), function(i)
      fmt_row(type[i], col[i], comp, comp$w_norm), character(1)),
    within_arm = vapply(seq_len(n()), function(i) {
      if (type[i] == "header") return("")
      vals <- vapply(arm_order, function(a) {
        da <- comp[comp$group_numeric == a, ]
        est_num(type[i], da[[col[i]]], da$w_norm)
      }, numeric(1))
      sprintf("%.1f-%.1f", min(vals), max(vals))
    }, character(1))
  )

arm_n  <- vapply(arm_order, function(a) sum(comp$group_numeric == a), integer(1))
table1 <- build %>%
  transmute(Variable = label,
            `Baseline (initial randomized sample)` = baseline,
            `Weighted follow-up`                   = weighted_fu,
            `Weighted within-arm range`            = within_arm) %>%
  bind_rows(tibble(
    Variable = "n",
    `Baseline (initial randomized sample)` = format(nrow(t1),   big.mark = ","),
    `Weighted follow-up`                   = format(nrow(comp), big.mark = ","),
    `Weighted within-arm range`            = sprintf("%d-%d", min(arm_n), max(arm_n))))

print(table1, n = Inf)
write.csv(table1,
          file.path(TABLES_DIR, "table1_demographics.csv"),
          row.names = FALSE, fileEncoding = "UTF-8")
```

Per-arm weighted follow-up table (one column per arm) for the appendix.

```{r table1-by-arm}
appendix <- tibble(Variable = items$label)
for (j in seq_along(arm_order)) {
  a  <- arm_order[j]
  da <- comp[comp$group_numeric == a, ]
  appendix[[arm_names[j]]] <- vapply(seq_len(nrow(items)), function(i) {
    if (items$type[i] == "header") return("")
    fmt_row(items$type[i], items$col[i], da, da$w_norm)
  }, character(1))
}
appendix <- bind_rows(appendix, as_tibble_row(
  c(Variable = "n", setNames(as.character(arm_n), arm_names))))

write.csv(appendix,
          file.path(TABLES_DIR, "supp_table_demographics_by_arm.csv"),
          row.names = FALSE, fileEncoding = "UTF-8")
```

# Differential attrition by cluster

Attrition by LPA cluster (assignments cover all N = 7,505 randomized).

```{r attrition-by-cluster}
attr_tbl <- mydata %>%
  mutate(cluster_assignment = factor(cluster_assignment,
                                     levels = c("1", "2", "3", "4"))) %>%
  group_by(cluster_assignment) %>%
  summarise(
    n_randomized = dplyr::n(),
    n_completers = sum(fup_complete == 1),
    n_missing    = sum(fup_complete == 0),
    pct_missing  = 100 * n_missing / n_randomized,
    .groups = "drop"
  )
print(attr_tbl)

# Overall attrition rate, for context
overall_pct <- 100 * sum(mydata$fup_complete == 0) / nrow(mydata)
cat(sprintf("\nOverall attrition: %.2f%% (%d of %d)\n",
            overall_pct,
            sum(mydata$fup_complete == 0),
            nrow(mydata)))

# Chi-square test: cluster x missingness
chisq_tab <- table(cluster = mydata$cluster_assignment,
                   missing = 1 - mydata$fup_complete)
print(chisq_tab)
chisq_res <- chisq.test(chisq_tab)
print(chisq_res)

# Logistic regression: P(missing) ~ cluster, Cluster 1 as reference
mydata$missing_fu <- as.integer(mydata$fup_complete == 0)
logit_fit <- glm(missing_fu ~ factor(cluster_assignment),
                 data = mydata, family = binomial())
logit_summary <- broom::tidy(logit_fit, exponentiate = TRUE,
                             conf.int = TRUE)
print(logit_summary)

write.csv(attr_tbl,
          file.path(TABLES_DIR, "supp_attrition_by_cluster.csv"),
          row.names = FALSE)
```

# Session info

```{r session-info}
sessionInfo()
```
